---
title: "Exercise Title Goes Here"
author: "Your Name Goes Here"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

<!-- These are global options  
       set root directory here 
       NOTE mine is active; yours is commented out
       change the path to reflect your CPU,
       then comment out or delete mine
--> 
```{r global_options, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/14842/Documents/SDHM/sdmR_ALLversions/sdhmR-V2022.1")
#knitr::opts_knit$set(root.dir = "~/sdhmR-V2022.1")
#knitr::opts_chunk$set(warning=FALSE,error=TRUE,message=FALSE)
```

---

## This exercise links to Module 2.5

Submit the completed exercise to us at:

* t.edwards@usu.edu 
* eric_tsakiris@fws.gov  

---

## Context

The data are of the *Pinus edulis* Common piñon, the species you dealt with in Exercise #3 -- #4.  This exercise relies on the pseudo--absence values you generated in exercise #4, and then subsequently added to the data presence values dataframe.  Hopefully you saved those combined presence:absence dataframe, or regrettably you will need to return to exercise #4 and re--build that dataframe.

The goal is to use the dataframe consisting of the presence:absence [X,Y]'s from exercise #4, and "punch" those [X,Y]'s through a stack of environmental predictor variables.  The extracted values are to be added to the exercise #4 dataframe, creating a new dataframe consisting of the: 

* FNETID ID, 
* the [X,Y]'s in the correct CRS, 
* the presence:absence coded as [0=absence, 1=presence], and 
* the predictor values associated with [X,Y] in the dataframe.

Remember, you will have been assigned data for *edulis* having one the following four discrete labels:

* **Seedlings**:  these data indicate spatial locations where seedlings of the tree have been found
* **Mortality**: these data indicate spatial locations where mortality has been observed
* **Persistence**: these data have no observed mortality or seedling at given spatial locations
* **Range**: the total dataset, including all spatial locations of seedlings, mortality, and persistence

---

## The Data

* Common piñon *Pinus edulis* sppcode=106
* The pseudo-absence data: Wherever stored from Homework #4
* Predictors: **`~\sdhmR-V2022.1\data\exercise\preds`** 
  * There are a total of 16 predictor variables the directory above
  * The variables are **`.img`** files
* Data files:
  * **`spp106pr_MORT.csv`**, **`spp106pr_PERS.csv`**, **`spp106pr_RANG.csv`**, **`spp106pr_SEED.csv`**

---

## Question #1

* Import the predictor GIS variables into R

```{r}
library(raster)
path.root <- "C:/Users/14842/Documents/SDHM/sdmR_ALLversions/sdhmR-V2022.1"
path.pred <- paste(path.root, "/data/exercise/preds", sep = "")

```


---

## Question #2

* Extract predictor variables from all 16 **`.img`** files using the presence:absence [X,Y]'s

```{r}
etpt5 <- raster("C:/Users/14842/Documents/SDHM/sdhmR-V2022.1/data/exercise/preds/etpt_5.img")
etpt6 <- raster("C:/Users/14842/Documents/SDHM/sdhmR-V2022.1/data/exercise/preds/etpt_6.img")

```

---

## Question #3

* Create a new dataframe by binding the extracted predictor variable values to the dataframe from exercise #4.

---

## Question #4

* Save your data as R objects:
  * Dataframe;
  * Point shapefile with geometry in R
  * Export as a point shapefile in ESRI format
* Save these R objects in a **`.RData`** file

These data will be used in the next exercise #6.

---

## The End

---
